Type Ia Supernova Explosions from Hybrid Carbon-Oxygen-Neon White Dwarf Progenitors

### 2019-02-09 paper/ms.tex
Even more on Carlyn's paper.

The simulations of supernova explosions presented here were performed
with a customized version of the Flash code, originally developed
at the University of Chicago.
Flash is a parallel, adaptive mesh, multi-physics simulation code
developed first for nuclear astrophysics applications and subsequently
for high-energy-density applications~\cite{Fryxetal00,calder.curtis.ea:high-performance,
calder.fryxell.ea:on,flash_pragmatic,flash_evolution}. 
Flash has been applied to a variety of astropysical problems by a host
of researchers, and the version we apply differs from other versions
principally in the modules describing thermonuclear burning. 
specifically routines to describe
thermonuclear burning during both the deflagration and detonation
phases, as well as routines to describe the evolution of
the dynamic ash.

The need for a model flame in simulating thermonuclear supernovae
follows from the scales of the problem. At high densities, the
width of a laminar nuclear flame is $< 1\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$
while the radius of the white dwarf 
$\sim 10^9\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$.
Even with adaptive mesh refinement, whole-star simulations cannot simultaneousely
directly simulate the nuclear flame, thus requiring use of a model to
describe the burning on unresovable scales. For the deflagration phase,
our burning model propagates an artificially
broadened flame with an advection-diffusion-reaction (ADR)
scheme~\cite{Khok95,VladWeirRyzh06} via evolution of
a reaction progress variable to describe the consumption of C
and additional variables to describe the evolution of intermediate-mass
elements into the stastical quasi-equilibruim of the 
Si-group~\cite{ifk1981,khok1981,khok1983} and then into iron-group elements
(IGEs) in full nuclear statistical equilibrium (NSE).

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel and $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP
 described by~\cite{VladWeirRyzh06}, 
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\cite{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning~\cite{timmes92,Chamulak2007The-Laminar-Fla},
these are boosted to account for speed-up of burning due to
unresolved buoyancy and background
turbulence~\cite{Khok95,gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.

>>>>

The simulations of supernova explosions presented here were performed
with a customized version of the Flash code, originally developed
at the University of Chicago.
Flash is a parallel, adaptive mesh, multi-physics simulation code
developed first for nuclear astrophysics applications and subsequently
for high-energy-density applications~\cite{Fryxetal00,calder.curtis.ea:high-performance,
calder.fryxell.ea:on,flash_pragmatic,flash_evolution}. 
Flash has been applied to a variety of astropysical problems by a host
of researchers, and the version we apply differs from other versions
principally in the modules describing thermonuclear burning via a 
flame capturing model.

The need for a model flame in simulating thermonuclear supernovae follows 
from the scales of the problem. At high densities, the width of a laminar 
nuclear flame is $< 1\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$ while the 
radius of the white dwarf $\sim 10^9\ensuremath{\;}{\ensuremath{\mathrm{cm}}}$.
Even with adaptive mesh refinement, whole-star simulations cannot 
simultaneousely resolve the nuclear flame, so simulations require a model to 
describe the burning on unresovable scales.  The model we apply is a flame 
capturing scheme and thermally-activated burning module to describe
thermonuclear burning during both the deflagration and detonation phases, 
as well as routines to describe the evolution of the dynamic ash. 
This description of the burning was developed during the course of research in 
thermonuclear supernovae and has been presented in a series of 
papers \citep[See][and references therein]{chandlery}. For completeness, 
we briefly review the flame capturing scheme here.

For the deflagration phase, the flame capturing scheme propagates an 
artificially broadened flame with an advection-diffusion-reaction (ADR)
scheme~\cite{Khok95,VladWeirRyzh06} via evolution of
a reaction progress variable to describe the consumption of C
and additional variables to describe the evolution of intermediate-mass
elements into the stastical quasi-equilibruim of the 
Si-group~\cite{ifk1981,khok1981,khok1983} and then into iron-group elements
(IGEs) in full nuclear statistical equilibrium (NSE).

The reaction progress variable is $\phi$ and it varies from $\phi=0$ for
unburned fuel and $\phi=1$ for burned ash. $\phi$ is evolved
via an advection-diffusion-reaction equation,
\begin{equation}
  \label{eq:ard}
  \partial_t \pv + \vec{u}\cdot\nabla \pv = \kappa \nabla^2 \pv +
\frac{1}{\tau} R\left(\phi\right) ,
\end{equation}
where $\vec{u}$ is the velocity of the fluid, $\kappa$ is the
diffusion coefficient, $\tau$ is the reaction timescale, and $R(\phi)$ is
a non-dimensional function describing the reaction. The parameters
$\kappa$, $\tau$, and $R(\phi)$ are tuned to propagate the reaction
front at a prescribed speed.  Our model uses the ``sharpened KPP
 described by~\cite{VladWeirRyzh06}, 
with $R\propto(\phi-\epsilon)(1-\phi+\epsilon)$, where
$\epsilon \simeq 10^{-3}$.  This scheme has been shown to be
acoustically quiet, stable, and to give a unique flame speed
\cite{townsley.calder.ea:flame}. The input flame speeds come from
tabulated results obtained by direct numerical simulations of
thermonuclear burning~\cite{timmes92,Chamulak2007The-Laminar-Fla},
these are boosted to account for speed-up of burning due to unresolved 
buoyancy and background turbulence~\cite{Khok95,
gamezo.khokhlov.ea:thermonuclear,townsley.calder.ea:flame,jacketal2014}.
